The Space of Feasible Execution Times for Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-09-26 dit-paper_ieee.tex
Shortened the article length (goal : strictly less than 7 pages) (Issue #18) Fixed #20 by adding a reference to the initial and final number of constraints in most cases

	The \emph{C-space} of a real-time system is, intuitively,
	the space of WCET values for which this system is feasible
	\cite{bini2004schedulability,george2009characterization}. A precise and efficient description of the C-space of a system would allow us to efficiently decide whether or not it is feasible on a set of candidate platforms.

@@

		Furthermore, the following notations are used:
		\begin{itemize}
			\item $\tau = \{\tau_1, \cdots, \tau_{n}\}$ is called the task system
			\item $H = \operatorname{lcm}(T_1, \cdots, T_{n})$
			\item $O_{max} = \max (O_1, \cdots, O_{n})$
			\item An \emph{idle time} is an instant $t$ at which each job that arrived
			strictly before $t$ has finished its execution.
			\item A \emph{busy period} is a time interval between two consecutive idle times in
			which at least one job is executing.
			\item The interval between the initial time ($t=0$) and the first
			idle time after $t=0$ is called the \emph{first busy period}.
		\end{itemize}

@@

  \begin{definition}
    The earliest DIT (if it exists) occurring in the system strictly after $t=0$ is called the
    \textbf{first DIT} and is noted $t_d$.
  \end{definition}

		\begin{remark} (from the definitions, if a DIT exist then):
			\begin{itemize}
				\item Contrary to other idle times, a DIT is independent of the
					scheduling or the execution times of the jobs.
				\item In a feasible system, every DIT is an idle time.
				\item In a feasible system, the first DIT happens after the end of the
					first busy period.
				\item In synchronous systems any $kH$ with $k \geqslant 0$ is a DIT and we
					also have $t_d \leqslant H$.
			\end{itemize}
		\end{remark}

		The value of the first DIT can be found by solving the following modular
		equations system:
		\begin{equation}
			\begin{split}
				 t_d = \min(& \{ t \; s.t. \; t > 0 \; and \\
						& \forall \tau_i \in \tau, \; \exists a_i \in [D_i, T_i]: t \equiv a_i \;
						(mod \; T_i) \})
			\end{split}
		\end{equation}
  The authors of \cite{george2009characterization} give a method to solve this system for cases in which periods are pairwise coprime.

@@

	\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figs/nofpdit_4.png}
	\end{center}
	\caption{Evolution of the existence of FPDIT with the CDF}
	\label{fig:noFPDIT}
	\end{figure}

@@

Such a linear problem can be solved by e.g. the simplex algorithm and a branch-and-cut approach. However, in our case the problem is not to test for redundancy of a particular constraint but to find all redundant constraints within a set. A naive method would be to test every constraint against all the others. We present a similar method with an efficient preprocessing when most constraints are redundant in Algorithm \ref{alg:prunRedun}.

@@

After doing this, no redundant constraints are left. However, this algorithm
takes an exponential time in the number of constraints. An efficient initial
description of $CSPACE$, as given by the methods described in previous
sections, is therefore preferred.

To summarize, we first reduced the number of constraints from every possible
interval $[t_1, t_2]$ to the intervals where $t_1$ is an arrival time and
$t_2$ is a deadline in $[t_d, t_d + H]$ if $t_d$ exists, or in
$[O_{max}, O_{max} + 2H]$ otherwise. This set of constraints is a
correct description of the C-Space, but it is not efficient. Redundant
constraints are removed through a linear integer approach.

@@

Let us now consider the same system but with a synchronous pattern of arrival.

    \begin{center}
    \begin{tabular}{|r|c|c|c|c|}
     \hline
      & $O_i$ & $C_i$ & $D_i$ & $T_i$ \\
     \hline
     $\tau_1$ & 0 & $C_1$ & 7 & 15\\
     \hline
     $\tau_2$ & 0 & $C_2$ & 2 & 5\\
     \hline
    \end{tabular}
    \end{center}

The first DIT happens at time $t_d = 7$. The intervals to consider are thus of the form $[0, t]$ where $t$ is a deadline happening before or at time $t_d$. This yields two intervals ($[0, 2]$ and $[0, 7]$) and the following constraints describing the C-space, none of them being redundant:
$$
\left\{
  \begin{array}{ccccc}
    & & C_2 & \leqslant & 2 \\
    C_1 & + & 2 C_2 & \leqslant & 7
  \end{array}
\right.
$$
This time, the C-space only contains 8 points: $(1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$.

@@

\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.5\textwidth]{figs/cspace_example.png}
  \caption{C-space of an asynchronous system and its corresponding synchronous system}
  \label{fig:cspaceComp}
\end{center}
\end{figure}

@@

    % PRECISER : on ne genere pas de systyemes avec instants synchroes

@@

	In other words, fig.~\ref{fig:sizeRatio}
	evaluates the feasibility gain of adding random offsets w.r.t. a synchronous task
	set. We have chosen in this paper to generate offsets randomly but we did not consider offsets that eventually lead to synchronous releases. The problem of offset assignment has been considered for example in \cite{grenier2008pushing}. The problem of those heuristics is that they depend on WCETs. Finding offsets that maximize a C-space region is still an open problem.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=0.5\textwidth]{figs/sizeratio.png}
			\caption{C-space size ratio between asynchronous and equivalent synchronous
			systems}
			\label{fig:sizeRatio}
		\end{center}
	\end{figure}

	The ratio between the size of the synchronous and asynchronous C-spaces of
	these systems roughly range from 0.95 in the worst case (high utilization, high
	CDF) to 0.85 in the best case (low utilization, low CDF). Each point on the graph is an average over 5000 tests.
	Note that the value of the CDF has a greater influence on this ratio than
	the system's total utilization, but less so when the CDF is higher. It is also
	important to note that even though a 0.95 ratio may not seem much, the
	synchronous C-spaces of the systems contain many dominated WCET vectors, and
	only those that are in the border (i.e. non-dominated) actually matter to the
	designer of the system. Using the asynchronous C-space instead will relax some
	constraints and therefore add new acceptable solutions that will dominate solutions that were previously non-dominated,
	increasing the quality of the C-space border.

@@

Future works could also repeat the simulation of section~\ref{sct:expCspaceGain} with the algorithm described in \cite{grenier2008pushing}, which determine offset values with optimal feasibility but for specific WCETs. We would like to study if those heuristics obtained for particular WCETs scenarios can be also interesting for any WCET in the C-Space.

% A mettre quelque part : nombre de contraintes avant/apres dans les expériences (issue #20)

>>>>

	The \mbox{C-space} of a real-time system is, intuitively,
	the space of WCET values for which this system is feasible
	\cite{bini2004schedulability,george2009characterization}. A precise and efficient description of the C-space of a system would allow us to efficiently decide whether or not it is feasible on a set of candidate platforms.

@@

		Furthermore, we define:
		\begin{itemize}
			\item $\tau = \{\tau_1, \cdots, \tau_{n}\}$ is called the task system
			\item $H = \operatorname{lcm}(T_1, \cdots, T_{n})$
			\item $O_{max} = \max (O_1, \cdots, O_{n})$
			\item An \emph{idle time} is an instant $t$ at which each job that arrived
			strictly before $t$ has finished its execution.
			\item A \emph{busy period} is a time interval between two consecutive idle times in
			which at least one job is executing.
			\item The interval between the initial time ($t=0$) and the first
			idle time after $t=0$ is called the \emph{first busy period}.
		\end{itemize}

@@

  \begin{definition}
    The earliest DIT (if it exists) occurring in the system strictly after $t=0$ is called the
    \textbf{first DIT}, denoted $t_d$.
  \end{definition}

		\begin{remark} (from previous definitions):
			\begin{itemize}
				\item Contrary to other idle times, DITs are independent of the
					scheduling or the execution times of the jobs.
				\item In a feasible system, every DIT is an idle time.
				\item In a feasible system, the first DIT happens after the end of the
					first busy period if it exists.
				\item In synchronous systems any $kH$ with $k \geqslant 0$ is a DIT and we
					also have $t_d \leqslant H$.
			\end{itemize}
		\end{remark}

		% The value of the first DIT can be found by solving the following modular
		% equations system:
		% \begin{equation}
		% 	\begin{split}
		% 		 t_d = \min(& \{ t \; s.t. \; t > 0 \; and \\
		% 				& \forall \tau_i \in \tau, \; \exists a_i \in [D_i, T_i]: t \equiv a_i \;
		% 				(mod \; T_i) \})
		% 	\end{split}
		% \end{equation}
  The authors of \cite{george2009characterization} give a method to find the value of $t_d$ for cases in which periods are pairwise coprime.

@@

	\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.4\textwidth]{figs/nofpdit_4.png}
	\end{center}
	\caption{Evolution of the existence of FPDIT with the CDF}
	\label{fig:noFPDIT}
	\end{figure}

@@

Such a linear problem can be solved by a branch-and-bound approach \cite{nemhauser1988integer}. However, in our case the problem is not to test for redundancy of a particular constraint but to find all redundant constraints within a set. A naive algorithm would consist of testing every constraint against all others. We present Algorithm \ref{alg:prunRedun}, which has the same theoretical complexity but is more efficient when most of the constraints are redundant. Indeed, most of the system we generated had a final C-space description size of less than five constraints, even if the initial description often contained more than a thousand.

@@

This algorithm
has an exponential complexity in the number of constraints. The initial description
of the C-space should thus be as concise as possible.

To summarize, we first reduced the number of constraints from every possible
interval $[t_1, t_2]$ to the intervals where $t_1$ is an arrival time and
$t_2$ is a deadline in $[t_d, t_d + H]$ if $t_d$ exists, or in
$[O_{max}, O_{max} + 2H]$ otherwise. This set of constraints is a
correct but redundant description of the C-Space. Redundant
constraints are removed through a linear integer approach.

@@

Let us now consider the same system but with a synchronous pattern of arrival ($O_1 = O_2 = 0$). The first DIT happens at time $t_d = 7$. The intervals to consider are thus of the form $[0, t]$ where $t$ is a deadline happening before or at time $t_d$. This yields two intervals ($[0, 2]$ and $[0, 7]$) and the following constraints describing the C-space, none of them being redundant:
$$
\left\{
  \begin{array}{ccccc}
    & & C_2 & \leqslant & 2 \\
    C_1 & + & 2 C_2 & \leqslant & 7
  \end{array}
\right.
$$
This time, the C-space only contains 8 points: $(1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$.

@@

\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.3\textwidth]{figs/cspace_example.png}
  \caption{C-space of an asynchronous system and its corresponding synchronous system}
  \label{fig:cspaceComp}
\end{center}
\end{figure}

@@

@@

	In other words, fig.~\ref{fig:sizeRatio}
	evaluates the feasibility gain of adding random offsets w.r.t. a synchronous task
	set. Offsets eventually leading to synchronous releases have been discarded. The problem of offset assignment has been considered for example in \cite{grenier2008pushing} in the case of fixed WCET values. To the best of our knowledge, finding offsets maximizing a C-space region is an open problem.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=0.4\textwidth]{figs/sizeratio.png}
			\caption{C-space size ratio between asynchronous and equivalent synchronous
			systems}
			\label{fig:sizeRatio}
		\end{center}
	\end{figure}

	The ratio between the size of the synchronous and asynchronous C-spaces of
	these systems roughly range from 0.95 in the worst case (high utilization, high
	CDF) to 0.85 in the best case (low utilization, low CDF). Each point on the graph is an average over 5000 tests.
	Note that the value of the CDF has a greater influence on this ratio than
	the system's total utilization, but less so when the CDF is higher. It is also
	important to note that even though a 0.95 ratio may not seem much, the
	synchronous C-spaces of the systems contain many dominated WCET vectors, and
	only those that are in the border (i.e. non-dominated) actually matter to the
	designer of the system. Using the asynchronous C-space instead will relax some
	constraints and therefore add new acceptable and dominating solutions,
	increasing the quality of the C-space border.

@@

Future works could also repeat the simulation of section~\ref{sct:expCspaceGain} with the algorithm described in \cite{grenier2008pushing}, which determine offset values with optimal feasibility for given WCET values. We could like to study if those heuristics obtained for particular WCETs scenarios can also be interesting for any WCET in the \mbox{C-Space}.
